---
title: "An area frame with points, a list frame, and their joint design"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An area frame with points, a list frame, and their joint design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6.5, fig.height = 4.5, dpi = 96)
```

National agricultural surveys increasingly combine an **area frame** (a grid
of square cells classified by intensity of agricultural use from satellite
imagery, with sample points inside the selected cells identifying the fields
and so the establishments) and a **list frame** (large and specialised
establishments, cheap to reach). This vignette walks through such a design
with `fieldopt`: stratified selection of cells, points inside them, the
multiplicity estimator, the cost of the field work, and the dual-frame and
two-stage allocations that put the pieces together. The data are synthetic;
the structure follows the design of a national survey with 1 km² cells.

## The area frame: cells with intensity classes

```{r frame}
library(fieldopt)
set.seed(2026)
side <- 30
cells <- expand.grid(x = seq_len(side), y = seq_len(side))
cells$unit <- paste0("c", seq_len(nrow(cells)))
use <- 0.5 * cells$x / side + 0.3 * sin(cells$y / 4) + rnorm(nrow(cells), sd = 0.15)
cells$intensity <- cut(use, c(-Inf, 0.25, 0.5, Inf), c("low", "mid", "high"))
cells$size <- c(low = 1, mid = 2, high = 4)[as.character(cells$intensity)]
table(cells$intensity)
```

Each cell holds a few establishments; the agricultural area of an
establishment inside a cell is what the point sampling needs. Here every
cell has two establishments with 40 % and 20 % of its area, and a crop value
per establishment that follows the intensity.

```{r establishments}
areas <- rbind(
  data.frame(establishment = paste0(cells$unit, "-1"), unit = cells$unit, area = 0.4),
  data.frame(establishment = paste0(cells$unit, "-2"), unit = cells$unit, area = 0.2))
areas$y <- round(rgamma(nrow(areas), shape = 4, rate = 4 / (20 * areas$area * cells$size[match(areas$unit, cells$unit)])), 1)
truth <- sum(areas$y)
truth
```

## Stratified selection of cells and points inside them

Cells are selected within intensity strata, more where use is intense, with
the local pivotal method so that the sample spreads over the territory; then
3, 6 or 9 points per cell depending on the stratum.

```{r select}
n_h <- c(low = 20, mid = 35, high = 45)
cells_s <- select_units(cells, n = n_h, strata = "intensity", seed = 7)
cells_s
ppc <- c(low = 3, mid = 6, high = 9)
points <- select_points(cells_s, points_per_cell = ppc, cell_size = 1, seed = 7)
nrow(points)
```

In the field each point is visited and the establishment it falls on is
interviewed about its whole operation. Here the field is simulated: a point
hits establishment 1 with probability 0.4, establishment 2 with probability
0.2, and otherwise falls on non-agricultural land.

```{r field}
set.seed(8)
u <- runif(nrow(points))
which_e <- ifelse(u < 0.4, "-1", ifelse(u < 0.6, "-2", NA))
hits <- data.frame(unit = points$unit, establishment = paste0(points$unit, which_e))[!is.na(which_e), ]
hits$y <- areas$y[match(hits$establishment, areas$establishment)]
nrow(hits); length(unique(hits$establishment))
```

## The multiplicity estimator

An establishment can be hit several times; the estimator divides the value of
every hit by the expected number of hits of that establishment under the
design, which depends on the cells it spans, their inclusion probabilities
and their point densities.

```{r estimate}
eh <- expected_hits(areas, cells_s, points_per_cell = ppc, cell_size = 1)
hits$expected_hits <- eh$expected_hits[match(hits$establishment, eh$establishment)]
point_estimator(hits, cells_s)
truth
```

The variance uses the cell design (local-mean estimator on the cell
contributions). The same cells selected as a spatially ordered systematic
sample with four interpenetrating replicates give a replicate-based variance
instead:

```{r systematic}
cells_r <- select_units(cells, n = n_h, strata = "intensity", method = "systematic", replicates = 4, seed = 7)
cells_r$crop <- tapply(areas$y, areas$unit, sum)[cells_r$unit]
design_variance(cells_r, "crop")
```

## What the field work costs

The cells are visited from a depot; the routing gives the travel, the cost
model the money. The cost per visited cell is the input the allocations
below need, and it depends on how many cells are visited and how spread they
are, which is why it is computed rather than assumed.

```{r cost}
frame <- rbind(data.frame(unit = "depot", x = side / 2, y = side / 2), cells[, c("unit", "x", "y")])
m <- travel_matrix(frame, method = "euclidean", unit = "km")
model <- field_cost_model(per_travel = 2.5, per_unit = 60, per_interview = 25,
                          interviews_per_unit = 4, currency = "BRL")
routes <- route_fieldwork(m, units = cells_s$unit[cells_s$sampled], depot = "depot",
                          max_stops = 8, cost_model = model)
routes$cost
c1 <- routes$cost[["total"]] / sum(cells_s$sampled)
c1
```

## Two-stage allocation: how many cells, how many points

With the between-cell and within-cell variance components (from a pilot or
from the frame) and the costs per cell and per point, Cochran's rule gives
the number of points per cell and then the number of cells for a target
precision.

```{r twostage}
cell_tot <- tapply(areas$y, areas$unit, sum)
two_stage_allocation(n_primary = nrow(cells), m_secondary = 9,
                     s2_between = var(cell_tot / 9), s2_within = var(areas$y) * 2,
                     c1 = routes$cost[["travel"]] / sum(cells_s$sampled) + model$per_unit,
                     c2 = model$per_interview, target_cv = 0.05, mean = mean(cell_tot) / 9)
```

## Dual-frame allocation: area plus list

The list frame holds the large establishments, which are also on the area
frame (the overlap `ab`); small establishments are on the area frame only
(`a`); a few specialised units are on the list only (`b`). Given the sizes,
means and standard deviations of the three domains and the cost per unit in
each frame (the area cost from the routing above), Hartley's allocation
gives the two sample sizes and the mixing weight, or the screening design
in which the area sample does not use the overlap.

```{r dualframe}
domains <- data.frame(domain = c("a", "ab", "b"),
                      size = c(1500, 250, 50),
                      mean = c(6, 30, 45), sd = c(5, 20, 35))
dual_frame_allocation(domains, cost_a = c1, cost_b = 45, deff_a = 1.6, target_cv = 0.05)
dual_frame_allocation(domains, cost_a = c1, cost_b = 45, deff_a = 1.6, theta = "screening", target_cv = 0.05)
```

`deff_a` carries the design effect of the point sampling relative to simple
random sampling of establishments; estimate it from the variance ratio of a
pilot. The allocation is a planning tool: the cost per area unit changes with
`n_A`, so run the routing at the allocated size and iterate once.

## References

Cochran, W. G. (1977). *Sampling Techniques*, 3rd ed. Wiley.

Hartley, H. O. (1974). Multiple frame methodology and selected applications.
*Sankhya C*, 36, 99--118.

Lohr, S. L. and Rao, J. N. K. (2006). Estimation in multiple-frame surveys.
*Journal of the American Statistical Association*, 101(475), 1019--1030.

Nealon, J. P. (1984). Review of the multiple and area frame estimators. USDA
Statistical Reporting Service, Staff Report 80.
