---
title: "An area frame with points, a list frame, and their joint design"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An area frame with points, a list frame, and their joint design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6.5, fig.height = 4.5, dpi = 96)
```

Agricultural surveys often combine an **area frame** (a grid of square cells
classified by intensity of agricultural use from satellite imagery, with
sample points inside the selected cells identifying the fields and so the
establishments) and a **list frame** (large and specialised establishments,
cheap to reach). This vignette walks through such a design with `fieldopt`:
stratified selection of cells, points inside them, the multiplicity
estimator, the cost of the field work, and the dual-frame and two-stage
allocations that put the pieces together. The data are synthetic; the
structure is the usual one of area frames built on a regular grid (square
segments of about 1 km²) with a list of large producers alongside.

## The area frame: cells with intensity classes

```{r frame}
library(fieldopt)
set.seed(2026)
side <- 30
cells <- expand.grid(x = seq_len(side), y = seq_len(side))
cells$unit <- paste0("c", seq_len(nrow(cells)))
use <- 0.5 * cells$x / side + 0.3 * sin(cells$y / 4) + rnorm(nrow(cells), sd = 0.15)
cells$intensity <- cut(use, c(-Inf, 0.25, 0.5, Inf), c("low", "mid", "high"))
cells$size <- c(low = 1, mid = 2, high = 4)[as.character(cells$intensity)]
table(cells$intensity)
```

Each cell holds a few establishments; the agricultural area of an
establishment inside a cell is what the point sampling needs. Here every
cell has two establishments with 40 % and 20 % of its area, and a crop value
per establishment that follows the intensity.

```{r establishments}
areas <- rbind(
  data.frame(establishment = paste0(cells$unit, "-1"), unit = cells$unit, area = 0.4),
  data.frame(establishment = paste0(cells$unit, "-2"), unit = cells$unit, area = 0.2))
areas$y <- round(rgamma(nrow(areas), shape = 4, rate = 4 / (20 * areas$area * cells$size[match(areas$unit, cells$unit)])), 1)
truth <- sum(areas$y)
truth
```

## Stratified selection of cells and points inside them

Cells are selected within intensity strata, more where use is intense, with
the local pivotal method so that the sample spreads over the territory; then
3, 6 or 9 points per cell depending on the stratum, laid out as a regular
grid with one random offset per cell (`layout = "systematic"`), which keeps
each point uniform over the cell while spreading them.

```{r select}
n_h <- c(low = 20, mid = 35, high = 45)
cells_s <- select_units(cells, n = n_h, strata = "intensity", seed = 7)
cells_s
ppc <- c(low = 3, mid = 6, high = 9)
points <- select_points(cells_s, points_per_cell = ppc, cell_size = 1, layout = "systematic", seed = 7)
nrow(points)
```

In the field each point is visited and the establishment it falls on is
interviewed about its whole operation. Here the field is simulated: a point
hits establishment 1 with probability 0.4, establishment 2 with probability
0.2, and otherwise falls on non-agricultural land.

```{r field}
set.seed(8)
u <- runif(nrow(points))
which_e <- ifelse(u < 0.4, "-1", ifelse(u < 0.6, "-2", NA))
hits <- data.frame(unit = points$unit, establishment = paste0(points$unit, which_e))[!is.na(which_e), ]
hits$y <- areas$y[match(hits$establishment, areas$establishment)]
nrow(hits); length(unique(hits$establishment))
```

## The multiplicity estimator

An establishment can be hit several times; the estimator divides the value of
every hit by the expected number of hits of that establishment under the
design, which depends on the cells it spans, their inclusion probabilities
and their point densities.

```{r estimate}
eh <- expected_hits(areas, cells_s, points_per_cell = ppc, cell_size = 1)
hits$expected_hits <- eh$expected_hits[match(hits$establishment, eh$establishment)]
point_estimator(hits, cells_s)
truth
```

The variance uses the cell design (local-mean estimator on the cell
contributions). The same cells selected as a spatially ordered systematic
sample with four interpenetrating replicates give a replicate-based variance
instead:

```{r systematic}
cells_r <- select_units(cells, n = n_h, strata = "intensity", method = "systematic", replicates = 4, seed = 7)
cells_r$crop <- tapply(areas$y, areas$unit, sum)[cells_r$unit]
design_variance(cells_r, "crop")
```

## What the field work costs

The cells are visited from a depot; the routing gives the travel, the cost
model the money. The cost per visited cell is the input the allocations
below need, and it depends on how many cells are visited and how spread they
are, which is why it is computed rather than assumed.

```{r cost}
frame <- rbind(data.frame(unit = "depot", x = side / 2, y = side / 2), cells[, c("unit", "x", "y")])
model <- field_cost_model(per_travel = 2.5, per_unit = 60, per_interview = 25,
                          interviews_per_unit = 4, currency = "BRL")
routed_unit_cost(frame, "depot", model, n = sum(n_h), method = "euclidean",
                 n_rep = 3, iterations = 60)
```

## The design effect of point sampling

The dual-frame allocation below needs the design effect of the area sample
relative to simple random sampling of establishments. `point_design_effect()`
simulates the whole area design (cells, points, hits, multiplicity
estimator) and returns the bias, the empirical variance, the mean of the
variance estimates, the number of interviews and the design effect.

```{r deff}
deff <- point_design_effect(cells, areas, n = n_h, strata = "intensity",
                            points_per_cell = ppc, cell_size = 1, n_sim = 60, seed = 9)
deff
```

The bias is negligible and the variance estimator of `point_estimator()`
tracks the empirical variance; the design effect is what enters `deff_a`.

## Two-stage allocation: how many cells, how many points

With the between-cell and within-cell variance components (from a pilot or
from the frame) and the costs per cell and per point, Cochran's rule gives
the number of points per cell and then the number of cells for a target
precision. `two_stage_design()` measures the cost per cell by routing a
sample of the allocated size and iterates until the number of cells settles.

```{r twostage}
cell_tot <- tapply(areas$y, areas$unit, sum)
ts <- two_stage_design(frame, "depot", model, m_secondary = 9,
                       s2_between = var(cell_tot / 9), s2_within = var(areas$y) * 2,
                       target_cv = 0.05, mean = mean(cell_tot) / 9,
                       method = "euclidean", n_rep = 3, iterations = 60)
ts
ts$history
```

## Dual-frame allocation: area plus list

The list frame holds the large establishments, which are also on the area
frame (the overlap `ab`); small establishments are on the area frame only
(`a`); a few specialised units are on the list only (`b`). Given the sizes,
means and standard deviations of the three domains, the cost per list unit
and the design effect from the simulation, `dual_frame_design()` runs
Hartley's allocation with the area cost measured by routing, iterating until
the area sample size settles; `theta = "screening"` gives the design in
which the area sample does not use the overlap.

```{r dualframe}
domains <- data.frame(domain = c("a", "ab", "b"),
                      size = c(1500, 250, 50),
                      mean = c(6, 30, 45), sd = c(5, 20, 35))
df <- dual_frame_design(frame, "depot", model, domains, cost_b = 45, deff_a = deff$deff,
                        target_cv = 0.05, method = "euclidean", n_rep = 3, iterations = 60)
df
df$history
tidy(df)
dual_frame_allocation(domains, cost_a = df$cost_a, cost_b = 45, deff_a = deff$deff,
                      theta = "screening", target_cv = 0.05)
```

## Estimating from both frames

Once both samples are in, `dual_frame_estimator()` combines them. Each
sampled unit carries its domain (`a` or `ab` on the area frame, `b` or `ab`
on the list); Hartley's estimator mixes the two estimates of the overlap with
a weight `theta` that can be fixed, set to zero (screening) or estimated to
minimise the variance, and the Fuller-Burmeister estimator adds the
difference between the two estimates of the overlap size. The variances come
from the variance estimator of each design.

```{r dual-estimate}
cells$domain <- ifelse(cells$intensity == "high" & runif(nrow(cells)) < 0.6, "ab", "a")
cells$crop_est <- ifelse(cells$domain == "ab", 60, 12) + rnorm(nrow(cells), sd = 3)
ab <- cells[cells$domain == "ab", ]   # the overlap: the same holdings, seen from the list
list_frame <- data.frame(unit = c(paste0("l", seq_len(nrow(ab))), paste0("b", 1:12)),
                         x = runif(nrow(ab) + 12), y = runif(nrow(ab) + 12),
                         domain = rep(c("ab", "b"), c(nrow(ab), 12)),
                         crop_est = c(ab$crop_est, 90 + rnorm(12, sd = 5)))
sa <- select_units(cells, n = 40, strata = "intensity", seed = 3)
sb <- select_units(list_frame, n = 20, method = "srs", seed = 4)
dual_frame_estimator(sa, sb, y = "crop_est", domain = "domain")
dual_frame_estimator(sa, sb, y = "crop_est", domain = "domain", theta = "screening")
dual_frame_estimator(sa, sb, y = "crop_est", domain = "domain", estimator = "fuller-burmeister")
```

## Auxiliaries: the ratio estimator and balanced samples

Area frames come with auxiliaries known for every cell: the area of farmland
from the land-cover map, the number of holdings from the last census. The
ratio estimator uses one such total at the estimation stage, and the cube
method uses several at the selection stage, drawing a sample whose
Horvitz-Thompson estimates of the auxiliaries match their totals.
`spatial_balance()` measures how evenly a sample covers the territory
(Stevens and Olsen's Voronoi measure: zero is perfect).

```{r auxiliaries}
cells$farmland <- cells$size * runif(nrow(cells), 15, 35)
cells$crop_r <- 0.5 * cells$farmland + rnorm(nrow(cells), sd = 3)
s_lpm <- select_units(cells, n = 40, strata = "intensity", seed = 5)
ratio_estimator(s_lpm, y = "crop_r", x = "farmland")[, c("total", "se", "variance", "variance_ht")]
s_cube <- select_units(cells, n = 40, strata = "intensity", method = "cube", balance = "farmland", seed = 5)
c(lpm = spatial_balance(s_lpm), cube = spatial_balance(s_cube),
  srs = spatial_balance(select_units(cells, n = 40, strata = "intensity", method = "srs", seed = 5)))
```

## References

Cochran, W. G. (1977). *Sampling Techniques*, third edition. Wiley.

Hartley, H. O. (1974). Multiple frame methodology and selected applications.
*Sankhya C*, 36, 99--118.

Lohr, S. L. and Rao, J. N. K. (2006). Estimation in multiple-frame surveys.
*Journal of the American Statistical Association*, 101(475), 1019--1030.

Nealon, J. P. (1984). Review of the multiple and area frame estimators. USDA
Statistical Reporting Service, Staff Report 80.

Deville, J.-C. and Tillé, Y. (2004). Efficient balanced sampling: the cube
method. *Biometrika*, 91(4), 893--912.

Fuller, W. A. and Burmeister, L. F. (1972). Estimators for samples selected
from two overlapping frames. *Proceedings of the Social Statistics Section,
American Statistical Association*, 245--249.

Stevens, D. L. and Olsen, A. R. (2004). Spatially balanced sampling of
natural resources. *Journal of the American Statistical Association*, 99,
262--278.
