---
title: "Routing for one or several teams"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Routing for one or several teams}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6.5, fig.height = 4.5, dpi = 96)
```

The same sample can be worked by one team on a long tour or by several teams
on short routes. The total travel differs, the number of days differs and the
cost differs. This vignette compares those organisations on one sample.

```{r setup}
library(fieldopt)
set.seed(7)
frame <- data.frame(unit = c("depot", paste0("s", 1:60)),
                    lat = c(-8.05, -8.05 + runif(60, -0.25, 0.25)),
                    lon = c(-35.00, -35.00 + runif(60, -0.35, 0.35)))
m <- travel_matrix(frame)
model <- field_cost_model(per_travel = 1.8, per_unit = 40, per_interview = 12,
                          interviews_per_unit = 6, currency = "BRL")
s <- select_units(frame[-1, ], n = 18, seed = 3)
sel <- s$unit[s$sampled]
```

## One tour

```{r one}
one <- route_fieldwork(m, units = sel, depot = "depot", cost_model = model)
one
```

## Routes limited by length

A team that must return to the depot every day has a maximum route length,
here 100 km. The solver splits the tour optimally under that limit (Prins
2004) and improves each route locally.

```{r days}
days <- route_fieldwork(m, units = sel, depot = "depot", max_length = 100,
                        cost_model = model)
days
autoplot(days)
```

## Routes limited by stops

A team that can interview at most four units a day has a stop limit instead.

```{r stops}
stops <- route_fieldwork(m, units = sel, depot = "depot", max_stops = 4,
                         cost_model = model)
stops
```

## Comparing

```{r compare}
data.frame(organisation = c("one tour", "100 km per route", "4 stops per route"),
           routes = c(one$n_routes, days$n_routes, stops$n_routes),
           travel_km = round(c(one$total, days$total, stops$total), 1),
           cost = round(c(one$cost[["total"]], days$cost[["total"]], stops$cost[["total"]])))
```

The travel grows with the number of routes because every route starts and
ends at the depot; the question for the planner is whether the extra kilometres
are cheaper than the extra days or teams, which is a cost model question, not
a routing one.

## Road distances

Distances from coordinates understate travel on real roads. Two ways to do
better, both feeding the same routing and cost functions:

* a **detour factor**, the ratio of road to straight-line distance, which on
  rural networks usually lies between 1.2 and 1.4: a planning estimate when
  no network is at hand;
* a **road-network matrix** from an OSRM server, with `method = "osrm"`:
  durations in minutes or distances in kilometres, requested in blocks so
  that the server's table limit is respected. The public demo server is for
  small tests; run your own server (with OpenStreetMap data of the region)
  for a frame. A matrix computed by any other engine can be passed through
  `matrix` instead.

```{r detour}
m_detour <- travel_matrix(frame, detour = 1.3)
m_detour
```

```{r network, eval = FALSE}
m_road <- travel_matrix(frame, method = "osrm", measure = "duration")
route_fieldwork(m_road, units = sel, depot = "depot", max_length = 480)   # 8 hours per route
```

The `snap` attribute of an OSRM matrix gives the distance from each unit to
the nearest point of the network the server could use; units snapped from
far away are the ones whose access the field team should check.

## Time at the unit, cost per team-day and the calendar

When the matrix is in minutes, a day has a length and so does the work at
each unit: `service_time` is added to the travel of a route when the limit
is checked (a segment with nine points and a few interviews can take two
hours), and the cost model's `per_route` charges each route as a team-day
(allowances, vehicle, lodging). Routes are then days of work.

```{r service}
m_min <- travel_matrix(frame, detour = 1.3, unit = "min")   # treat the km as minutes at 60 km/h
daily <- field_cost_model(per_travel = 0.5, per_route = 400, per_unit = 20, per_interview = 15,
                          interviews_per_unit = 3, currency = "BRL")
day <- route_fieldwork(m_min, units = sel, depot = "depot", max_length = 480, service_time = 120,
                       iterations = 50, cost_model = daily)
day
glance(day)
```

With several teams based in several towns, `schedule_fieldwork()` assigns
the units to the nearest base (or balances them when a base is short of
team-days), routes each base and distributes the routes over the teams and
the days available, longest first. The calendar says who goes where on which
day and whether the work fits.

```{r schedule}
frame2 <- rbind(frame, data.frame(unit = "base2", lat = -8.30, lon = -35.30))
m2 <- travel_matrix(frame2, detour = 1.3, unit = "min")
sch <- schedule_fieldwork(m2, units = sel, teams = c(depot = 1, base2 = 1), days = 3,
                          max_length = 480, service_time = 120, iterations = 50, cost_model = daily)
sch
head(sch$calendar[, c("base", "team", "day", "stops", "travel", "duration")])
autoplot(sch)
```

## Solver settings

`iterations` is the number of GRASP restarts; `alpha` the greediness of the
construction (0 is pure greedy, 1 pure random). The default 200 iterations
with `alpha = 0.3` is enough for a few hundred units; the `gap` field reports
how far the solution is from a two-edge lower bound, which is loose, so gaps
of 20 to 40 percent are normal for good solutions.

```{r settings}
quick <- route_fieldwork(m, units = sel, depot = "depot", iterations = 10, seed = 2)
long <- route_fieldwork(m, units = sel, depot = "depot", iterations = 500, seed = 2)
c(quick = quick$total, long = long$total)
```
