---
title: "Routing for one or several teams"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Routing for one or several teams}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6.5, fig.height = 4.5, dpi = 96)
```

The same sample can be worked by one team on a long tour or by several teams
on short routes. The total travel differs, the number of days differs and the
cost differs. This vignette compares those organisations on one sample.

```{r setup}
library(fieldopt)
set.seed(7)
frame <- data.frame(unit = c("depot", paste0("s", 1:60)),
                    lat = c(-8.05, -8.05 + runif(60, -0.25, 0.25)),
                    lon = c(-35.00, -35.00 + runif(60, -0.35, 0.35)))
m <- travel_matrix(frame)
model <- field_cost_model(per_travel = 1.8, per_unit = 40, per_interview = 12,
                          interviews_per_unit = 6, currency = "BRL")
s <- select_units(frame[-1, ], n = 18, seed = 3)
sel <- s$unit[s$sampled]
```

## One tour

```{r one}
one <- route_fieldwork(m, units = sel, depot = "depot", cost_model = model)
one
```

## Routes limited by length

A team that must return to the depot every day has a maximum route length,
here 100 km. The solver splits the tour optimally under that limit (Prins
2004) and improves each route locally.

```{r days}
days <- route_fieldwork(m, units = sel, depot = "depot", max_length = 100,
                        cost_model = model)
days
autoplot(days)
```

## Routes limited by stops

A team that can interview at most four units a day has a stop limit instead.

```{r stops}
stops <- route_fieldwork(m, units = sel, depot = "depot", max_stops = 4,
                         cost_model = model)
stops
```

## Comparing

```{r compare}
data.frame(organisation = c("one tour", "100 km per route", "4 stops per route"),
           routes = c(one$n_routes, days$n_routes, stops$n_routes),
           travel_km = round(c(one$total, days$total, stops$total), 1),
           cost = round(c(one$cost[["total"]], days$cost[["total"]], stops$cost[["total"]])))
```

The travel grows with the number of routes because every route starts and
ends at the depot; the question for the planner is whether the extra kilometres
are cheaper than the extra days or teams, which is a cost model question, not
a routing one.

## A road-network matrix

Distances from coordinates understate travel on real roads. When a routing
engine is available, pass its matrix: the units must be in the same order as
the coordinates, which are then used only for names and plots.

```{r network, eval = FALSE}
# with the osrm package and a running server
net <- osrm::osrmTable(src = frame[, c("unit", "lon", "lat")],
                       dst = frame[, c("unit", "lon", "lat")])$durations
m_net <- travel_matrix(frame, matrix = net, unit = "min")
route_fieldwork(m_net, units = sel, depot = "depot", max_length = 480)
```

## Solver settings

`iterations` is the number of GRASP restarts; `alpha` the greediness of the
construction (0 is pure greedy, 1 pure random). The default 200 iterations
with `alpha = 0.3` is enough for a few hundred units; the `gap` field reports
how far the solution is from a two-edge lower bound, which is loose, so gaps
of 20 to 40 percent are normal for good solutions.

```{r settings}
quick <- route_fieldwork(m, units = sel, depot = "depot", iterations = 10, seed = 2)
long <- route_fieldwork(m, units = sel, depot = "depot", iterations = 500, seed = 2)
c(quick = quick$total, long = long$total)
```
